Specific leaf mass explains difference in biomass partitioning between woody plant functional types

### 2014-10-14 manuscript/draftpaper.Rnw
Added least-square means of LMF and LAR. Now all analyses use a version of the dataset where leaf mass and area are included (to avoid using different datasets).

\item We used linear mixed-effects models (with species within study as the random effect) to test for significance and to calculate confidence intervals for  PFT-specific averages.

@@

\item We found similar results for the above-ground leaf mass fraction (LMF), where a clear and strong separation was found between plant functional types (Fig.~\ref{fig:figure3}, see also Fig.~\ref{fig:figureSI5}). For the leaf area ratio (LAR), PFTs were generally closer together, with a notable exception for small evergreen gymnosperms (< 1m height) (Fig.~\ref{fig:figure4}).

\item Results for variance partitioning of LMF and LAR gave comparable results (Table~\ref{varpartpipe}): PFT increased the variance explained in LMF from \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[2],2)} to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[3],2)}\%. However, SLM was not a good substitute for PFT, as the explained variance decreased to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[4],2)}\%. The leaf area ratio, similarly to $A_F/A_S$, did not depend strongly on PFT or SLM.

@@

>>>>

\item We used linear mixed-effects models (with species within study as the random effect) to test for significance and to calculate confidence intervals for PFT-specific averages. Least-square means for LMF and LAR were calculated with the lsmeans package.

@@

\item We found similar results for the above-ground leaf mass fraction (LMF), where a clear and strong separation was found between plant functional types (Fig.~\ref{fig:figure3}, see also Fig.~\ref{fig:figureSI5}). For the leaf area ratio (LAR), PFTs were generally closer together, with a notable exception for small evergreen gymnosperms (< 1m height) (Fig.~\ref{fig:figure4}). Calculated least-square means of LMF and LAR (averages accounting for variation explained by plant height) showed very similar patterns: large differences in mass-based allocation between PFTs (LMF), and small differences in area-based allocation (LAR) (Fig.~\ref{fig:figure5}).

\item We also partionined the variance of LMF and LAR (Table~\ref{varpartpipe}). When added to a model that included plant height and biome, PFT increased the variance explained in log-transformed LMF from \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[2],2)} to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[3],2)}\%. However, SLM was not a good substitute for PFT, as the explained variance decreased to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[4],2)}\%. For (log-transformed) LAR, PFT and SLM explained very little additional variance.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{figure5_lsmeans_LMFLAR.pdf}
    \caption{Average leaf mass fraction (LMF) and leaf area ratio (LAR) accounting for variation explained by plant height (least-square means) with a mixed effects model (with species within study as a random effect). Error bars are 95\% confidence intervals.}
    \label{fig:figure5}
\end{figure}
